Future challenges for Research Software Engineering

### 2025-04-14 future_challenges_for_RSEng.tex
fix up sentence.

% Article frontmatter
% ===================
\title{Future challenges for Research Software Engineering} % Title of the article
%\short{} % Short title of the article (optional)
\author{
Florian Goth\texorpdfstring{\autref{1}}{},
Leyla Jael Castro\texorpdfstring{\autref{1}}{},
Gerasimos Chourdakis\texorpdfstring{\autref{1}}{},
Simon Christ\texorpdfstring{\autref{1}}{},
Jeremy Cohen\texorpdfstring{\autref{1}}{},
Jean-Noël Grad\texorpdfstring{\autref{1}}{},
Magnus Hagdorn\texorpdfstring{\autref{1}}{},
Toby Hodges\texorpdfstring{\autref{1}}{},
Jan Linxweiler\texorpdfstring{\autref{1}}{},
Frank Löffler\texorpdfstring{\autref{1}}{},
Michele Martone\texorpdfstring{\autref{1}}{},
Jan Philipp Thiele\texorpdfstring{\autref{1}}{},
Nick del Grosso\texorpdfstring{\autref{1}}{}
} % Authors and references to addresses
\institute{\autlabel{1} Fantasy University} % Institutes with labels
\abstract{Admittedly, it’s still only 2025 and we are actually stillworking on establishing Research Software Engineering
as a new domain with the hope of ensuring that it gets the recognition it deserves.
And while it is certainly too early to already herald the end of RSEng,
we should acknowledge that the future is in constant flux and
it is worthwhile to discuss how RSEng will/could change in response to these upcoming challenges in an open discussion.} % Abstract of the article
\keywords{Keywords go here.} % Keywords for the article

@@

\subsection{Ethics and social consideration}
The ethics and values of the Research Software Engineer are grounded in their practice.
They commit to the health, safety and welfare of the public and act in the interest of society, their employer and their clients \cite{Goth2023}.
Sometimes these values are in direct conflict with the aims of their institution.
For example, their PI might insist on quickly developing some software without adhering to current best practices.
Another example might be that they are developing code that can also be used for military purposes, such as numerical solvers or computer vision systems.
Like engineers, RSEs take the responsibility for the software they develop.
RSEs need to be aware of these issues and develop their own ``red lines''.

RSEs are publicly funded.
Similarly to how data collected by scientists should adhere to the FAIR principles \cite{FAIR}, research software developed by RSEs should adhere to the FAIR4RS principles~\cite{FAIR4RS}.

A code of ethics and professional conduct for RSEs should be developed by institutions together with RSEs themselves.
This can then form the basis for institutional policies and guidelines.

\begin{FramedParagraphWithFootnotes}  
FG-NOTE: Given what ChatGPT has put in there, I wonder if we need to separate these two items
MM-NOTE: (Flo believes it was from Michele) Are we assuming that progress will bring prosperity? Will we remain constrained by needing to produce products/artifacts?

@@

- Corporate dominance:
  - AI and coding platforms might be increasingly controlled by a small number of large corporations.
  - RSE community might need to develop or advocate for open-source alternatives.
\end{FramedParagraphWithFootnotes}

>>>>

% Article frontmatter
% ===================
\title{Future challenges for Research Software Engineering} % Title of the article
%\short{} % Short title of the article (optional)
\author{
Florian Goth\texorpdfstring{\autref{1}}{},
Leyla Jael Castro\texorpdfstring{\autref{1}}{},
Gerasimos Chourdakis\texorpdfstring{\autref{1}}{},
Simon Christ\texorpdfstring{\autref{1}}{},
Jeremy Cohen\texorpdfstring{\autref{1}}{},
Jean-Noël Grad\texorpdfstring{\autref{1}}{},
Magnus Hagdorn\texorpdfstring{\autref{1}}{},
Toby Hodges\texorpdfstring{\autref{1}}{},
Jan Linxweiler\texorpdfstring{\autref{1}}{},
Frank Löffler\texorpdfstring{\autref{1}}{},
Michele Martone\texorpdfstring{\autref{1}}{},
Jan Philipp Thiele\texorpdfstring{\autref{1}}{},
Nick del Grosso\texorpdfstring{\autref{1}}{}
} % Authors and references to addresses
\institute{\autlabel{1} Fantasy University} % Institutes with labels
\abstract{It's 2025 and we are, admittedly, still working on establishing and growing Research Software Engineering (RSE)
as a domain in its own right, with the hope of ensuring that it gets the recognition it deserves.
Nonetheless, as the use of digital tools becomes ever more pervasive in a researchers day-to-day tasks, there are already people suggesting that RSE will
cease to exist as a separate field, and the specialist skills that RSEs provide to support and undertake research will
become a core part of the tooling and skill set of researchers in almost all domains.
Are we already approaching "the end of Research Software Engineering"?! While we believe that it is certainly too early to
already herald the demise of RSE as a separate field,
we also recognise the need to acknowledge that the future is in constant flux and
it is worthwhile discussing how RSE will/could change in response to these upcoming challenges in an open discussion.} % Abstract of the article
\keywords{Keywords go here.} % Keywords for the article

@@

\subsection{Ethics and social consideration}
FG-NOTE: Given what ChatGPT has put in there, I wonder if we need to separate these two items
MM-NOTE: (Flo believes it was from Michele) Are we assuming that progress will bring prosperity? Will we remain constrained by needing to produce products/artifacts?

@@

- Corporate dominance:
  - AI and coding platforms might be increasingly controlled by a small number of large corporations.
  - RSE community might need to develop or advocate for open-source alternatives.
